Characterisation of MSC in clinical trial reports: analysis of published descriptors

### 2020-06-04 paper2_figures.Rmd
HDI and JIF added to dataset
figures 1 - 6 complete

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  pdf_document: default
---

@@

# Figures specification
* titled
* cited in numerical order in the text using Arabic numbers. 
* no additional charges for color. 
* submitted as individual files
* labeled with the Corresponding Author name, the appropriate figure number, and  orientation (e.g., “top”)
* Figures may be submitted as multipart panels.
*  *.tif* or .eps 
* final publication size; 1-column width or, if necessary, 1½ column width. The  2-column width should not be used unless necessary.
* 1.0 column = 3.2 Inches, 8.2 Centimeters, 19.25 Picas
* 1.5 column = 4.5 Inches, 11.4 Centimeters, 27.00 Picas
* 2.0 column = 5.8 Inches, 14.8 Centimeters, 35.00 Picas
* height of all figures <= to 9.6inches / 24.5 cm / 58 picas.
* Calibri font sub sans serif Arial
* Multipanel labeled using uppercase 12-point Calibri Bold. 
* minimum text size of 6 points
* bold enough to be easily read after reduction, as should all symbols used in the * figure. Line or bar graphs or flow charts with text should be created in black and white, not shades of gray, which are difficult to reproduce in even tones. If more than two sets of data are represented, use of fill patterns or colors (not gray) is suggested to present the data clearly.
* Minimum resolution is 300 dpi for color and grayscale figures, 600 dpi for  combination halftones, and 1000 dpi for line art.

```{r pkg}
library(tidyverse)
# library(readxl)
# library(kableExtra)
library(tiff)
library(grid)
library(patchwork)
library(janitor)

@@

# figure saving settings
units = "in"
fig_w <- 3.2
fig_w_1.5 <- 4.5
fig_w_2 <- 5.8
fig_h <- 4.5
fig_h_1.5 <- 4.5 * 1.5
dpi <- 600
device <- "tiff"

@@

```{r}
file <- "data/checked_processed_aw_corrections.csv"
dat <- read_csv(file)
# write(names(dat), "data/varnames.txt")

@@

## Fig 1 - Article search funnel

```{r data-prep-fig1}
# create data
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

@@

# modify the labels
labels$step_names2 <- c( "EXCLUDE conference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology applied microbiology oncology\nhaematology, engineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative medicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT TYPE\narticle",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells OR\nMSC AND characterisation AND clinical")
```

```{r fig1}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               size = 0) +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 3,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP" ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 7, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = rev(pal))

ggsave("figs/fig_1_Article_search_funnel.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

## Figure 3 Articles by country

```{r country-prep}
summary_country <- dat %>%
  group_by(country) %>% 
  summarise(n = length(country))

@@

```{r fig3-country-bar}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3],
           size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 7, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```
## Figure 4 Articles by Phase

### Eight phase version

```{r phase-prep}

summary_phase <- dat %>%
  group_by(phase) %>% 
  summarise(n = length(phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

@@

@@

```{r fig4-phase}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(phase,n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3], size = 0) +
  xlab("") +
  geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 1),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

## Figure 5 Articles by Source-Year

```{r source_year_prep}
summary_tissue <- dat %>% 
  group_by(source, year, 
           .drop = FALSE) %>% 
  summarise(n = length(source))
dat %>% 
  group_by(source, 
           .drop = FALSE) %>% 
  summarise(n =length(source), pc = n/84)

@@

### stacked
```{r fig5-source-year-stack}
source_year <- summary_tissue %>% 
  ggplot(aes(x = year, y = n, fill = source)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal[c(1,3,5,6)],
                    guide = guide_legend(nrow = 3, title = NULL )) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 20, 2),
                    minor_breaks = seq(0, 20, 2),
                    limits = c(0, 20),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = c(0.2, 0.9),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))

ggsave("figs/fig_5_Article_source_year_stack.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

## Figure 6 - three panels

### panel 1: Articles by Test quality

```{r test-quality-prep}
summary_test <- dat %>% 
  group_by(test_results_2) %>% 
  summarise(n = length(test_results_2),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```{r testquality}
test_bar <- summary_test %>% 
  ggplot(aes(x = reorder(test_results_2, n), y = n, fill = test_results_2)) +
  geom_bar(stat = "identity", size = 0) +
  xlab("") +
  geom_text(aes(y = n + 6, label = pc) ) +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 50, 5),
                     minor_breaks = seq(0, 50, 1),
                     limits = c(0, 50),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0),
                   labels = c("Stated as 'standard\nphenotype' (s)",
                              "Phenotypic markers\nreported (y)",
                              "Tests done but\nnot reported (t)",
                              "Not discussed (o)",
                              "Average value (n)")) +
  labs(subtitle = "A") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4),
        plot.margin = margin(r = 0.1, b = 0.1, l = 0, unit="in")) +
  coord_flip()

# ggsave("figs/fig_6_Article_test_quality.tif", 
#        plot = test_bar, 
#        device = device,
#        width = fig_w, 
#        height = fig_w,
#        units = units,
#        dpi = dpi)

### panel 2 how many markers had average values  
```{r}
dat %>% 
  filter(test_results_2 != "Average value (n)") %>%
  filter(marker_count == 1) %>% 
  select(ref_number,test_results_2, test_results)

summary_marker_count <- dat %>% 
  filter(test_results_2 == "Average value (n)") %>%
  group_by(marker_count) %>% 
  summarise(n = length(marker_count))

```{r markercount}
marker_bar <- summary_marker_count %>% 
  ggplot(aes(x = factor(marker_count), y = n)) +
  geom_bar(stat = "identity", fill = pal[2], size = 0) +
  scale_y_continuous(name = "Number of\n'Average value (n)' Articles ", 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 12),
                     expand = c(0, 0)) +
  scale_x_discrete(name = "Number of markers\nreported by an article",
                   expand = c(0, 0)) +
  labs(subtitle = "B") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.margin = margin(r = 0.2, t = 0,l = 0, unit="in")) 
```

### panel 3  Articles by Phase and Test quality reported (previously fig7 a)

### Stacked by number of articles
```{r}

dat %>%
  group_by(phase_condensed) %>% 
  summarise(n = length(phase_condensed),
            pc = n/84)

summary_phase_test <- dat %>%
  group_by(phase_condensed, test_results_2) %>% 
  summarise(n = length(phase_condensed))
```

```{r phasetest}
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = phase_condensed, y = n, fill = test_results_2)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
 labs(subtitle = "C") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 8),
        legend.position = "none", 
        plot.margin = margin(r = 0.3, l = 0.1, t = 0, unit="in"))

```{r fig6}

layout <- "
aa
bc
"

fig6 <- test_bar + marker_bar + phase_test + 
  plot_layout(design = layout) +
  plot_annotation(title = "Wilson, Alison, Figure 6. TOP")&
    theme(text = element_text(size = 8, hjust = 0))

ggsave("figs/fig_6.tif", 
       plot = fig6, 
       device = device,
       width = fig_w_2, 
       height = fig_h,
       units = units,
       dpi = dpi)

## Figure 7 Characterisation by marker.
```{r}

@@

# change levels of status and markers.
# markers
# process test_col names
test_cols <- str_replace(test_cols, "_test", "")
# set the levels in that order
test_by_attribute$marker2 <- factor(charact_by_marker_ref$marker2, 
                                    levels = test_cols)

# status
# recode status
test_by_attribute <- test_by_attribute %>%
  mutate(test_status = recode_factor(status,
                                     p = "Performed, value reported",
                                     m = "Performed, no value reported",
                                     n = "Not mentioned"))

test_by_attribute$test_status <- fct_relevel(test_by_attribute$test_status,
                                             "Not mentioned",
                                             "Performed, no value reported",
                                             "Performed, value reported")

>>>>

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  svg_document: default
---

@@

# Figures specification
* titled
* cited in numerical order in the text using Arabic numbers. 
* no additional charges for color. 
* submitted as individual files
* labeled with the Corresponding Author name, the appropriate figure number, and  orientation (e.g., “top”)
* Figures may be submitted as multipart panels.
*  *.svg* or .eps 
* final publication size; 1-column width or, if necessary, 1½ column width. The  2-column width should not be used unless necessary.
* 1.0 column = 3.2 Inches, 8.2 Centimeters, 19.25 Picas
* 1.5 column = 4.5 Inches, 11.4 Centimeters, 27.00 Picas
* 2.0 column = 5.8 Inches, 14.8 Centimeters, 35.00 Picas
* height of all figures <= to 9.6inches / 24.5 cm / 58 picas.
* Calibri font sub sans serif Arial
* Multipanel labeled using uppercase 12-point Calibri Bold. 
* minimum text size of 6 points
* bold enough to be easily read after reduction, as should all symbols used in the * figure. Line or bar graphs or flow charts with text should be created in black and white, not shades of gray, which are difficult to reproduce in even tones. If more than two sets of data are represented, use of fill patterns or colors (not gray) is suggested to present the data clearly.
* Minimum resolution is 300 dpi for color and grayscale figures, 600 dpi for  combination halftones, and 1000 dpi for line art.

```{r pkg}
library(tidyverse)
# library(readxl)
# library(kableExtra)
library(grid)
library(patchwork)
library(janitor)

@@

# figure saving settings
units = "in"
fig_w <- 3.5
fig_w_1.5 <- 5
fig_w_2 <- 7.3
fig_h <- fig_w
device <- "svg"

@@

```{r}
file <- "data/checked_processed_aw_corrections.csv"
dat <- read_csv(file, skip = 1)
# write(names(dat), "data/varnames.txt")

@@

## Fig 1 - Search

### prep

```{r data-prep-fig1}
# create data for funnel 1A
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

@@

# modify the labels
labels$step_names2 <- c( "EXCLUDE,\nconference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology, applied\nmicrobiology, oncology\nhaematology,\nengineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative,\nmedicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT\nTYPE article",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells\nOR MSC AND characterisation,\nAND clinical")

# read data for 1B
fig1b <- read.table("data/fig1b.txt", header = TRUE)

@@

### each panel of fig 1
```{r fig1}
# figure 1 a
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               size = 0) +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 3,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 7, colour = "black", hjust = 0, face = "plain"),
        panel.spacing = unit(0, "lines"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
    scale_fill_manual(values = rev(pal))

# fig 1b

data_collection <- fig1b %>% 
  ggplot() +
  geom_text(aes(x = x, y = y, label = item),
            size = 2.5,nudge_y = 0.1,
            hjust = "left") +
  facet_grid(. ~ data_element) +
  scale_x_continuous(limits = c(0, 7)) +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(hjust = 0, size = 5.5, colour = "white", face = "bold"),
        strip.background = element_rect(fill = pal[1], colour = pal[1], size = 0.05),
        panel.border = element_rect(colour = "black", fill = NA),
        panel.spacing = unit(0, "lines"),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

### whole fig 1
```{r fig1}

layout <- "
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
bbbbbbbbbbbbbbbbbbbb
bbbbbbbbbbbbbbbbbbbb
bbbbbbbbbbbbbbbbbbbb
"

fig1 <- article_funnel + data_collection + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 1, face = "bold"))

ggsave("figs/fig_1.svg", 
       plot = fig1, 
       device = device,
       width = fig_w, 
       height = fig_h*1.6,
       units = units)

## Fig 2 - Articles by country, phase, indications, route of admin

### prep

```{r fig2-prep}
# by country A
summary_country <- dat %>%
  group_by(country) %>% 
  summarise(n = length(country))

# phase B
summary_phase <- dat %>%
  group_by(phase) %>% 
  summarise(n = length(phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

@@

# indication C
summary_indication <- dat %>%
  group_by(indication) %>% 
  summarise(n = length(indication),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# route D
summary_route <- dat %>%
  group_by(route) %>% 
  summarise(n = length(route),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

@@

### each panel figure 2
```{r fig2}
#by country
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[4],
           size = 0, width = 0.8) +
  xlab("") +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 15, 5),
                     minor_breaks = seq(0, 15, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by phase
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(phase,n), y = n)) +
  geom_bar(stat = "identity", fill = pal[4], size = 0, width = 0.8) +
  xlab("") +
 # geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 35, 10),
                     minor_breaks = seq(0, 35, 1),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by indication
indication_bar <- summary_indication %>% 
  ggplot(aes(x = reorder(indication, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[4], size = 0, width = 0.8) +
  xlab("") +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 13),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by route
route_bar <- summary_route %>% 
  ggplot(aes(x = reorder(route, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[4], size = 0, width = 0.8) +
  xlab("") +
#  geom_text(aes(y = n + 3, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 25, 5),
                     minor_breaks = seq(0, 25, 1),
                     limits = c(0, 25),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

@@

### whole fig 2
```{r fig2-layout}

layout <- "
a#
a#
ad
ad
ad
ad
ad
bd
bd
bd
bd
cd
cd
cd
"

fig2 <- country_bar  + route_bar+ phase_bar+ indication_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Number of Articles",
                  tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, face = "plain"))

ggsave("figs/fig_2.svg", 
       plot = fig2, 
       device = device,
       width = fig_w_2, 
       height = fig_h*2,
       units = units)

@@

## Fig 3 - Articles by sources, allo/auto stem/stromal

### prep

```{r fig3_prep}
# source 3a
summary_source <- dat %>% 
  group_by(source) %>% 
  summarise(n = length(source),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# allo va auto 3b
summary_allo_auto <- dat %>% 
  group_by(allo_auto) %>% 
  summarise(n = length(allo_auto),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# stem vs stromal 3b
summary_stem_stromal <- dat %>% 
  group_by(stem_stromal) %>% 
  summarise(n = length(stem_stromal),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
```
### each panel figure 2

```{r fig_3}
# source
source_bar <- summary_source %>% 
  ggplot(aes(x = source, y = n)) +
  geom_bar(stat = "identity",fill = pal[4], size = 0, width = 0.8) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black",
                                  angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())

# allo - auto
allo_auto_bar <- summary_allo_auto %>% 
  ggplot(aes(x = allo_auto, y = n)) +
  geom_bar(stat = "identity",fill = pal[4], size = 0, width = 0.8) +
  scale_y_continuous(name = NULL,
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black", 
                                  angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())

# stem_stomal
stem_stromal_bar <- summary_stem_stromal %>%
  ggplot(aes(x = stem_stromal, y = n)) +
  geom_bar(stat = "identity",fill = pal[4], size = 0, width = 0.8) +
  scale_y_continuous(name = NULL,
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black", 
                                   angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

### whole fig 3
```{r fig_3-layout}
layout <- "
aaabbbcccc
"

fig3 <- source_bar  + allo_auto_bar+ stem_stromal_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Source",
                  tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, vjust = 1, face = "plain"))

ggsave("figs/fig_3.svg", 
       plot = fig3, 
       device = device,
       width = fig_w_1.5, 
       height = fig_h,
       units = units)
```

## Fig 4 - characterisation by article

### prep

```{r fig4_prep}
# test stringency
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
# phase by stringency
summary_phase_test <- dat %>%
  group_by(phase_condensed, test_results) %>% 
  summarise(n = length(phase_condensed))

# number of markers reported in articles either reporting % by batch of average percentages (p)
summary_marker_p_count <- dat %>% 
  filter(test_results != "3. Tests done but % not reported") %>%
  filter(test_results != "4. Stated as 'standard phenotype'") %>%
  filter(test_results != "5. No characterisation discussed") %>%
  group_by(marker_p_count) %>% 
  summarise(n = length(marker_p_count))

```
### each panel figure 4

```{r fig_4}
# test stringency overall
test_bar <- summary_test %>% 
  ggplot(aes(x = test_results, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0) +
  xlab("") +
  geom_text(aes(y = n + 6, label = pc), size = 3 ) +
  scale_fill_manual(values = rev(pal[1:5]),
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 50, 5),
                     minor_breaks = seq(0, 50, 1),
                     limits = c(0, 53),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none") +
  coord_flip()

# tests stringency by phase
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0, width = 0.8) +
  scale_fill_manual(values = pal,
                    name = NULL) +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

# 
markerp_bar <- summary_marker_p_count %>% 
  ggplot(aes(x = marker_p_count, y = n)) +
  geom_bar(stat = "identity", fill = pal[6], size = 0, width = 0.8) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 12),
                     expand = c(0, 0)) +
  scale_x_continuous(name = "Number of markers",
                     breaks = seq(0, 11, 1),
                    minor_breaks = seq(0, 12, 2),
                    limits = c(0.5, 12),
                     expand = c(0, 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())

@@

### whole fig 4
```{r fig_4-layout}
layout <- "
#aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa
bbbcccccccccccccccccccccccccccccccccccc
"

layout <- "
aaaaaabbbcccc
"

fig4 <- test_bar + phase_test + markerp_bar +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))

ggsave("figs/fig_4.svg", 
       plot = fig4, 
       device = device,
       width = fig_w_2, 
       height = 2.2,
       units = units)
```

## Fig 5 - Characterisation detail

### prep

@@

# change levels of status and markers.
# markers
# process test_col names
test_cols <- str_replace(test_cols, "_test", "")
# set the levels in that order
test_by_attribute$marker2 <- factor(test_by_attribute$marker2, 
                                    levels = test_cols)
test_by_attribute$ref_number <- factor(test_by_attribute$ref_number, 
                                    levels = c(levels(test_by_attribute$ref_number), "Total P", "Total M"))
# status
# recode status
test_by_attribute <- test_by_attribute %>%
  mutate(test_status = recode_factor(status,
                                     p = "Performed, value reported (by batch or average)",
                                     m = "Performed, no value reported",
                                     n = "Not performed"))

test_by_attribute$test_status <- fct_relevel(test_by_attribute$test_status,
                                             "Not performed",
                                             "Performed, no value reported",
                                             "Performed, value reported (by batch or average)")
test_by_attribute <- test_by_attribute %>% 
  select(ref_number, marker2, test_status)

# test_by_attribute$n <- 1
